---
title: "bbmm_io"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Provides simple tools to reduce the output size of rasters with extensive **zero** and **NA** regions.


# The original data
First read in an example dataset (that is also the template) and plot it.

```{r orig}
source("io.R")
orig = read_template(mask = FALSE)
pretty_plot(orig, 
            title = "original", 
            axes = TRUE)
```


# Encoding
Now we encode it which generates a matrix for non-zero and non-NA pixels.  We get back `index`, `lon`, `lat` and `value`.

```{r encode}
e = encode_raster(orig)
head(e)
```

## Or just write it and ignore encoding

We can skip that step by simply calling the encoded writer.

```{r write}
e = write_encoded(orig, file = "encoded.bin")
```

# Read it back

Now read it back.

```{r read}
x = read_encoded(file = "encoded.bin")
pretty_plot(x,
            title = "post-encoding",
            axes = TRUE)
```

Well, the seem to be the same.

# Difference

What about the difference?  Say `x - orig`?

```{r diff}
d = x[1] - orig[1]
d[[1]] |>
  as.vector() |>
  summary()
```

# Storage demand

Let's see about the size of the files.

```{r sizes}
orig_size = file.info("Jul01.nc")$size
encoded_size = file.info("encoded.bin")$size

cat("original nc size:", orig_size, "\n")
cat("encoded bin size:", encoded_size, "\n")
cat("ratio original/encoded:", orig_size/encoded_size, "\n")
cat("\n\nOh\n\n")
```
